Hydra Battles and Cie

### 2021-08-25 doc/part-hydras.tex
Alectryon:  chapter 6 (completed, draft)

\begin{Coqsrc}
Lemma H'_phi0_omega : forall k, H'_ (phi0 omega) k =
                               iterate H'_succ_fun  k (H'_ omega) k.
Proof with auto with E0.
  intro k; rewrite H'_eq3, <- H'_phi0_Si ...
  -  rewrite CanonS_Canon, CanonS_phi0_lim;  f_equal ...
Qed.
\end{Coqsrc}

Using extensionality of the functional \texttt{iterate}, we can get a closed formula.

\begin{Coqsrc}
Lemma H'_phi0_omega_closed_formula k :
  H'_ (phi0 omega) k =
  iterate (fun (f: nat -> nat) (l : nat) => iterate  f (S l) l)
               k
               (fun k : nat => S (2 * k)%nat)
               k.
\end{Coqsrc}

Note that this short formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
and the inner one  first-order (on type \texttt{nat}).

@@

\begin{Coqsrc}
Remark H'_non_mono1 :
  ~ (forall alpha beta k, (alpha o<= beta)%e0 ->
                          (H'_ alpha k <= H'_ beta k)%nat).
Proof.
 intros H ;specialize (H 42 omega 3).
 assert (H0 :(42 o<= omega)%e0) by (repeat split; auto).  
 apply H in H0; rewrite H'_Fin, H'_omega  in H0; lia.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Section Proof_of_Abstract_Properties.
  Record P (alpha:E0) : Prop :=
    mkP {
        PA : strict_mono (H'_ alpha);
        PB : alpha <> Zero -> forall n,  (n < H'_ alpha n)%nat;
        PC : H'_ alpha <<= H'_ (Succ alpha);
        PD : dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha);
        PE : forall beta n, Canon_plus n alpha beta -> 
                            (H'_ beta n <= H'_ alpha n)%nat}.

Theorem P_alpha : forall alpha, P alpha.
  Proof.
    intro alpha; apply well_founded_induction with lt.
   (* rest of proof skipped *)

Section Proof_of_Abstract_Properties.
\end{Coqsrc}

By elimination, we get a catalogue of properties of the functions $L'_\alpha$:

\begin{Coqsrc}
Section Abstract_Properties.
 Variable alpha : E0.

 Theorem H'_alpha_mono : strict_mono (H'_ alpha).

 Theorem H'_alpha_gt : alpha <> Zero ->
        forall n, n < H'_ alpha n.

 Theorem H'_alpha_Succ_le :
        H'_ alpha <<= H'_ (Succ alpha).

 Theorem H'_alpha_dom :
    dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).

 Theorem H'_restricted_mono_l :
    forall beta n, Canon_plus n alpha beta -> 
                                 H'_ beta n <= H'_ alpha n.

  Lemma H'_alpha_ge_id : id <<= H'_ alpha.

  Lemma H'_alpha_mono_weak : forall k l, k <= l ->
                                        H'_ alpha k <= H'_ alpha l.

End Abstract_Properties.
\end{Coqsrc}

@@

\noindent
\begin{Coqsrc}
  H'_dom : forall alpha beta : E0, alpha o< beta ->
  dominates_strong (H'_ beta) (H'_ alpha).
\end{Coqsrc}

@@

\begin{Coqsrc}
 Theorem H'_L_ alpha: forall i:nat,  (H'_ alpha i <= L_ alpha (S i))%nat.
\end{Coqsrc}

@@

\noindent
\emph{From Module~\href{../theories/html/hydras.Epsilon0.L_alpha.html}{Epsilon0.L\_alpha}}

\begin{Coqsrc}
Theorem battle_length_std_Hardy (alpha : E0) :
  alpha <> Zero ->
  forall k , 1 <= k -> exists l: nat,  
       H'_ alpha k - k <= l /\
       battle_length standard k (iota (cnf alpha)) l.    
\end{Coqsrc}

@@

\index{hydras}{Library Prelude!iterate}
\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

@@

\index{coq}{Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  F_ (Canon alpha i)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

\begin{Coqanswer}
The command has indeed failed with message:
In environment
alpha : E0
notlimit : Limitb alpha = false
nonzero : alpha <> Zero
i : nat
F_ : forall x : E0, nat -> x o< alpha -> nat
The term "F_ (Pred alpha) ?x" has type "Pred alpha o< alpha -> nat"
while it is expected to have type 
"Pred alpha o< alpha -> Pred alpha o< alpha"
(cannot unify "nat" and "Pred alpha o< alpha").
\end{Coqanswer}

@@

\begin{Coqsrc}
Definition call_lt (c c' : E0 * nat) :=
  lexico Lt (Peano.lt) c c'.

Lemma call_lt_wf : well_founded call_lt.
  unfold call_lt; apply Inverse_Image.wf_inverse_image,  wf_lexico.
  -  apply E0.Lt_wf.
  -  unfold Peano.lt; apply Nat.lt_wf_0. 
Qed.

Instance WF : WellFounded call_lt := call_lt_wf.

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Limitb alpha) :=
                 { | left _ => F_star (Canon alpha i,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

(* Finally, F_ alpha is defined as its first iterate ! *)

Definition F_  alpha i := F_star (alpha, 1) i.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma F_zero_eqn : forall i, F_ Zero i = S i.

Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
                               F_ alpha i = F_ (Canon alpha i) i.

Lemma F_succ_eqn : forall alpha i,
    F_ (Succ alpha) i = iterate (F_ alpha) (S i) i.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma LF1 : forall i,  F_ 1 i = S (2 * i).

Lemma LF2 : forall i, exp2 i * i < F_ 2 i.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem F_alpha_mono alpha : strict_mono (F_ alpha).

Theorem F_alpha_ge_S alpha : forall n, n < F_ alpha n.

Theorem F_alpha_Succ_le alpha : F_ alpha <<= F_ (Succ alpha).

Theorem F_alpha_dom alpha : 
     dominates_from 1 (F_ (Succ alpha)) (F_ alpha).

Theorem F_restricted_mono_l alpha : 
   forall beta n, Canon_plus n alpha beta -> 
                                        F_ beta n <= F_ alpha n.
\end{Coqsrc}

As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}, which states that the function \texttt{F\_} is ``monotonous'' w.r.t. its first argument and ``domination''.

\begin{quote}
  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
\end{quote}

\begin{Coqsrc}
Lemma F_mono_l: forall alpha beta : E0, 
   beta o< alpha -> dominates (F_ alpha) (F_ beta).
\end{Coqsrc}

>>>>

\input{movies/snippets/Hprime/HprimePhi0Omega}

Using extensionality of the functional \texttt{iterate}, we also get a closed formula.

\input{movies/snippets/Hprime/HprimePhi0OmegaClosed}

Note that this formula contains two occurrences of the functional \texttt{iterate}, the outer one is in fact a second-order iteration (on type \texttt{nat -> nat)}
and the inner one  first-order (on type \texttt{nat}).

@@

\input{movies/snippets/Hprime/HprimeNonMono1}

@@

\input{movies/snippets/Hprime/PAP}

% By elimination, we get a catalogue of properties of the functions $L'_\alpha$:

% \begin{Coqsrc}
% Section Abstract_Properties.
%  Variable alpha : E0.

%  Theorem H'_alpha_mono : strict_mono (H'_ alpha).

%  Theorem H'_alpha_gt : alpha <> Zero ->
%         forall n, n < H'_ alpha n.

%  Theorem H'_alpha_Succ_le :
%         H'_ alpha <<= H'_ (Succ alpha).

%  Theorem H'_alpha_dom :
%     dominates_from 1 (H'_ (Succ alpha)) (H'_ alpha).

%  Theorem H'_restricted_mono_l :
%     forall beta n, Canon_plus n alpha beta -> 
%                                  H'_ beta n <= H'_ alpha n.

%   Lemma H'_alpha_ge_id : id <<= H'_ alpha.

%   Lemma H'_alpha_mono_weak : forall k l, k <= l ->
%                                         H'_ alpha k <= H'_ alpha l.

% End Abstract_Properties.
% \end{Coqsrc}

@@

\noindent
\input{movies/snippets/Hprime/HprimeDom}

@@

\input{movies/snippets/L_alpha/HprimeL}

@@

\noindent
\emph{From Module~\href{../theories/html/hydras.Hydra.Hydra_Theorems.html}{Hydra.Hydra\_Theorems}}

\input{movies/snippets/Hydra_Theorems/battleLengthStdHardy}

@@

\index{hydras}{Library Prelude!iterate}

@@

 \input{movies/snippets/F_alpha/FailDemo}

@@

\input{movies/snippets/F_alpha/goodDef}

@@

\input{movies/snippets/F_alpha/FEquations}

@@

\input{movies/snippets/F_alpha/FirstValues}

@@

\input{movies/snippets/F_alpha/FalphaThms}

As a corollary, we prove the following proposition, p. 284 of~\cite{KS81}.
\begin{quote}
  If $\beta<\alpha$, $F_\alpha$ dominates $F_\beta$.
\end{quote}

\input{movies/snippets/F_alpha/FDomContext}
\input{movies/snippets/F_alpha/FDom}
